\chapter{Scan conversion}
Also called rasterization, an old problem. Generation of discrete pixelvalues. Approximation by a finite number of pixels. 
\section{Bresenham line}
Choose the closest pixel at each intersection. 
\begin{compactdesc}
	\item[\lp{Goal}]
 Fast decision which pixel has to be drawn next. Criterion: Position of the midpoint $\vtr{m}$ with respect to the intersection point $\vtr{q}$.
\pgr{5-1_ScanConversion}{6}{-2.6}{-1.39}{-0.2}{0.4}
Implicit equation of the straight line
\begin{gather*}
	f(x,y)=ax+by+c=0\\
	d=f(\vtr{m})=f(x_p+1,y_p+1/2)
\end{gather*}
$d>0\Rightarrow$ Select pixel NE, $d<0\Rightarrow$ select pixel E.
\item[\lp{Update criterion}] 
Update criterion, for E:
\pgr{5-1_ScanConversion}{8}{-3.0}{-1.39}{-0.1}{0.7}
\begin{align*}
	d_{\text{old}}&=f(x_p+1,y_p+1/2)\\
	&=a(x_p+1)+b(y_p+1/2)+c\\
	d_{\text{new}}&=f(x_p+2,y_p+1/2)\\
	&=a(x_p+2)+b(y_p+1/2)+c
\end{align*}
\begin{gather*}
	d_{\text{new}}-d_{\text{old}}=a=\Delta y
\end{gather*}
Update criterion for NE.
\begin{align*}
	d_{\text{old}}&=f(x_p+1,y_p+1/2)\\
	&=a(x_p+1)+b(y_p+1/2)+c\\
	d_{\text{new}}&=f(x_p+2,y_p+3/2)\\
	&=a(x_p+2)+b(y_p+3/2)+c
\end{align*}
\begin{gather*}
	d_{\text{new}}-d_{\text{old}}=a+b=\Delta y-\Delta x
\end{gather*}
Filled polygons (especially triangles) are the most important graphics primitives. GPUs can process up to 50 millions of triangles/second.
\item[\lp{Spatial coherence}] Straightforward solution: Inside test for each pixel - inefficient! Instead: Process scan-line after scan-line. Span: Group of picked pixels inside a sca-line.
\item[\lp{Spans}]
Span 1: from a to b. Span 2: from c to d.
\item[\lp{Algorithm}]\hfill \\
\begin{enumerate*}[label=\protect\circled{\arabic*},itemjoin=]
	\item Calculate all intersections on the scan line\\
	\item Sort the intersection points by ascending x-coordinates.\\
	\item Fill all spans in between two consecutive intersection points if the parity is odd.
\end{enumerate*}
\pgr{5-1_ScanConversion}{15}{-1.3}{-1.32}{2.5}{0.7}
\end{compactdesc}
