\chapter{Scale-space representations}
From an original signal $f(x)$ generate a parametric family of signals $f^t(x)$, where fine-scale information is successively suppressed.
\section{Image pyramid}
\begin{inparaenum}[\itshape Level 1:]
	\setcounter{enumi}{-1}
	\item $1\times1$,
	\item $2\times2$,
	\item $4\times 4$,
	\item[Level $J-1$:] $N/2\times N/2$,
	\item[Level $J$ (base):] $N\times N$.
\end{inparaenum}
\section{Applications}
\begin{inparaenum}[\itshape(1)]
	\item Search for correspondence: look at coarse scales, then refine with finer scales
	\item Edge tracking: a ``good'' edge at a fine scale has parents at a coarser scale
	\item Control of detail and computational cost in matching: e.g. finding stripes; terribly important in texture representation
\end{inparaenum}
\section{Pyramids}
\begin{compactdesc}
	\item[\lp{Gaussian pyramid}] 
		Smooth with gaussians, because $\text{``$\text{gaussian}^2$''}=\text{another gaussian}$. 
		%\emph{Synthesis}: smooth and sample. \emph{Analysis}: take the top image. Gaussians are low pass filters, so representation is redunant.
		Progressively blurred and subsampled versions of the image. Adds scale invariance to fixed-size algorithms.
	\item[\lp{Laplacian Pyramid}]<++>
		%\emph{Synthesis}: Preserve difference between upsampled gaussian pyramid level and Gaussian pyramid level. Band pass filter - each leevl representns spatial frequencies (largely) unrepresented at other levels. \emph{Analysis}: Reconstruct Gaussian pyramid, take top layer.
		Shown the information added in gaussian pyramid at each spatial scale. Useful for noise reduction \& coding.
	\item[\lp{Wavelet/QMF}] Bandpassed representation, complete, but with aliasing and some non-oriented subbands. Recursive application of a two-band filter bank to the lowpass band of the provious stage yields octave band splitting.
	\item[\lp{Steerable pyramid}] Shown components at each scale and orientation separately, Non-aliased subbands. Good for texture and feature analysis.
		\section{Haar Transform}
		Two major sub-operations:
		\begin{inparaenum}[\itshape(1)]
			\item Scaling captures info at different frequencies
			\item Translation captures info at different locations
		\end{inparaenum}
		Can be represented by filtering and downsampling. Relatively poor energy compaction.
\end{compactdesc}
